---
title: "DeltaExposome_DeltaCognition"
output: html_document
date: '2023-11-29'
---

# Load required packages
```{r}
library(tidyr)
library(dplyr)
library(ggplot2)
library(lsr)
library(lme4)
library(lmerTest)
library(mediation)
library(ggcorrplot)
library(corrplot)
library(RColorBrewer)
library(glmnet)
library(mgcv)
library(gamm4)
library(readr)
library(devtools)
library(R.matlab)
library(mgcv)
library(ggplot2)
library(visreg)
library(ggtext)
library(sjPlot)
library(LaCroixColoR)
```



# Load in the bifactor scores
```{r}
bifactor_scores<-read.csv("/Users/askeller/Documents/ExposomeAnalyses/LongitudinalExpScores/abcd_longitudinal_factor_scores_7f_bifactor.csv")
colnames(bifactor_scores)[1]<-"subjectkey"
bifactor_scores_T0 <- bifactor_scores[bifactor_scores$Timepoint=="baseline_year_1_arm_1",]
bifactor_scores_T2 <- bifactor_scores[bifactor_scores$Timepoint=="2_year_follow_up_y_arm_1",]
colnames(bifactor_scores_T2)<-c("subjectkey","ID_Timepoint","Timepoint","General_SES_T2","School_T2","Family_Values_T2","Family_Turmoil_T2","Dense_Urban_Poverty_T2","Extracurriculars_T2","Screen_Time_T2","Timepoint_numeric","Site")
```



# Load in the cognition data
```{r}
nihtb<-read.table("/Users/askeller/Documents/Kellernet_PrelimAnalysis/NihTB/abcd_tbss01.txt",header=TRUE)

# Convert cognitive data to numeric values
nihtb$nihtbx_flanker_uncorrected <- as.numeric(nihtb$nihtbx_flanker_uncorrected)
nihtb$nihtbx_picvocab_uncorrected <- as.numeric(nihtb$nihtbx_picvocab_uncorrected)
nihtb$nihtbx_pattern_uncorrected <- as.numeric(nihtb$nihtbx_pattern_uncorrected)
nihtb$nihtbx_picture_uncorrected <- as.numeric(nihtb$nihtbx_picture_uncorrected)
nihtb$nihtbx_reading_uncorrected <- as.numeric(nihtb$nihtbx_reading_uncorrected)

# Extract and rename cognitive variables for baseline data and T2 data
nihtb_T0 <- nihtb[nihtb$eventname=="baseline_year_1_arm_1",c("subjectkey","nihtbx_picvocab_uncorrected","nihtbx_flanker_uncorrected","nihtbx_pattern_uncorrected","nihtbx_reading_uncorrected","nihtbx_picture_uncorrected")]
colnames(nihtb_T0)<-c("subjectkey","PicVocab","Flanker","Pattern","Reading","Picture")
nihtb_T2 <- nihtb[nihtb$eventname=="2_year_follow_up_y_arm_1",c("subjectkey","nihtbx_picvocab_uncorrected","nihtbx_flanker_uncorrected","nihtbx_pattern_uncorrected","nihtbx_reading_uncorrected","nihtbx_picture_uncorrected")]
colnames(nihtb_T2)<-c("subjectkey","PicVocab_T2","Flanker_T2","Pattern_T2","Reading_T2","Picture_T2")

```



# Load in Family and Site covariates
```{r}
family <-read.table("/Users/askeller/Documents/Kellernet_PrelimAnalysis/ses/acspsw03.txt",header=TRUE)
family_T0<-family[family$eventname=="baseline_year_1_arm_1",]
site_info <- readRDS("/Users/askeller/Documents/Kellernet_PrelimAnalysis/DEAP-siteID.rds")
site_T0<-site_info[site_info$event_name=="baseline_year_1_arm_1",]
colnames(site_T0)[1]<-"subjectkey"
```



# Combine the data into one dataframe
```{r}
step1 <- merge(nihtb_T0,bifactor_scores_T0[,c("subjectkey",setdiff(colnames(bifactor_scores_T0),colnames(nihtb_T0)))],by="subjectkey")
step2 <- merge(step1,family_T0[,c("subjectkey",setdiff(colnames(family_T0),colnames(step1)))],by="subjectkey")
step3 <- merge(step2,site_T0[,c("subjectkey",setdiff(colnames(site_T0),colnames(step2)))],by="subjectkey")
step5 <- merge(nihtb_T2,bifactor_scores_T2[,c("subjectkey",setdiff(colnames(bifactor_scores_T2),colnames(nihtb_T2)))],by="subjectkey")
step6 <- merge(step3,step5[,c("subjectkey",setdiff(colnames(step5),colnames(step3)))],by="subjectkey")

# this dataframe is n=6,530 with 69 variables
```


# Subset the participants with all data available for these analyses
```{r}
abcd.data<-step6[,c("subjectkey","interview_age","sex","abcd_site","rel_family_id","race.6level","household.income","high.educ","sex_at_birth","General_SES","School","Family_Values","Family_Turmoil","Dense_Urban_Poverty","Extracurriculars","Screen_Time","PicVocab","Flanker","Picture","Pattern","Reading","General_SES_T2","School_T2","Family_Values_T2","Family_Turmoil_T2","Dense_Urban_Poverty_T2","Extracurriculars_T2","Screen_Time_T2","PicVocab_T2","Flanker_T2","Picture_T2","Pattern_T2","Reading_T2")]

abcd.data<-abcd.data[complete.cases(abcd.data),]

```


# Make sure age is numeric
```{r}
abcd.data$interview_age<-as.numeric(abcd.data$interview_age)
```


# Calculate changes in Exposome and Cognition:
```{r}
# Delta cognition
abcd.data$delta_PicVocab <- abcd.data$PicVocab_T2 - abcd.data$PicVocab
abcd.data$delta_Flanker <- abcd.data$Flanker_T2 - abcd.data$Flanker
abcd.data$delta_Picture <- abcd.data$Picture_T2 - abcd.data$Picture
abcd.data$delta_Pattern <- abcd.data$Pattern_T2 - abcd.data$Pattern
abcd.data$delta_Reading <- abcd.data$Reading_T2 - abcd.data$Reading

# Delta exp-factor
abcd.data$delta_General_SES <- abcd.data$General_SES_T2 - abcd.data$General_SES
```


# adjust change scores by starting scores
```{r}
abcd.data$true_exposome_change <- lm(delta_General_SES~General_SES,data=abcd.data)$residuals
abcd.data$true_PicVocab_change <- lm(delta_PicVocab~PicVocab,data=abcd.data)$residuals
abcd.data$true_Flanker_change <- lm(delta_Flanker~Flanker,data=abcd.data)$residuals
abcd.data$true_Picture_change <- lm(delta_Picture~Picture,data=abcd.data)$residuals
abcd.data$true_Pattern_change <- lm(delta_Pattern~Pattern,data=abcd.data)$residuals
abcd.data$true_Reading_change <- lm(delta_Reading~Reading,data=abcd.data)$residuals
```


# z-score the change scores
```{r}
abcd.data$true_PicVocab_change<-(abcd.data$true_PicVocab_change-mean(abcd.data$true_PicVocab_change,na.rm=TRUE))/sd(abcd.data$true_PicVocab_change,na.rm=TRUE)
abcd.data$true_Flanker_change<-(abcd.data$true_Flanker_change-mean(abcd.data$true_Flanker_change,na.rm=TRUE))/sd(abcd.data$true_Flanker_change,na.rm=TRUE)
abcd.data$true_Picture_change<-(abcd.data$true_Picture_change-mean(abcd.data$true_Picture_change,na.rm=TRUE))/sd(abcd.data$true_Picture_change,na.rm=TRUE)
abcd.data$true_Pattern_change<-(abcd.data$true_Pattern_change-mean(abcd.data$true_Pattern_change,na.rm=TRUE))/sd(abcd.data$true_Pattern_change,na.rm=TRUE)
abcd.data$true_Reading_change<-(abcd.data$true_Reading_change-mean(abcd.data$true_Reading_change,na.rm=TRUE))/sd(abcd.data$true_Reading_change,na.rm=TRUE)
```


# linear models of change x change
```{r}
mod1<-lmer(true_PicVocab_change~interview_age+sex_at_birth+(1|rel_family_id)+(1|abcd_site)+true_exposome_change,data=abcd.data)
mod2<-lmer(true_Flanker_change~interview_age+sex_at_birth+(1|rel_family_id)+(1|abcd_site)+true_exposome_change,data=abcd.data)
mod3<-lmer(true_Picture_change~interview_age+sex_at_birth+(1|rel_family_id)+(1|abcd_site)+true_exposome_change,data=abcd.data)
mod4<-lmer(true_Pattern_change~interview_age+sex_at_birth+(1|rel_family_id)+(1|abcd_site)+true_exposome_change,data=abcd.data)
mod5<-lmer(true_Reading_change~interview_age+sex_at_birth+(1|rel_family_id)+(1|abcd_site)+true_exposome_change,data=abcd.data)


tab_model(mod1,mod2,mod3,mod4,mod5,pred.labels = c("Intercept","Age","Sex","Delta Exp-Factor"),dv.labels=c("Picture Vocabulary","Flanker","Pic. Sequence Mem.","Pattern Comparison","Reading Recognition"),digits=2,digits.p=2,show.se=TRUE,show.stat=TRUE,string.stat = "t",string.se = "Std. Error",show.intercept=TRUE,show.ci = FALSE, p.adjust="bonferroni",p.style="scientific", file = "/Users/askeller/Documents/ExposomeAnalyses/LongitudinalExpScores/DeltaExpFactor_DeltaCog_True_FullSample_z.doc")
```

